{\includegraphics[width=3cm, height=2.5cm]{Cran.jpg}} \\ \includegraphics[width=7cm, height=1cm]{TB.jpg} \\ {Technical Documentation for TersectBrowser+}

### 2025-05-05 TechDoc/TechDoc.tex
Formatting

\section{Architecture}
The TersectBrowser+ software is an Angular project, with an original frontend and backend written in Angular 8, and a separate extension space written in Angular 19 using Node 22. This 'bolt-on' approach was chosen for the extensions in order to seamlessly integrate the new extensions with the original workspace within the project time constrainsts. Additionally, this approach allows the new extnesions to use up-to-date packages, resulting in enhanced functionality and a modern design. To run TersectBrowser+ on different platforms, the new extensions are hosted on NPM and can be linked as dependencies within the project.

@@

The Admin is defined as an individual with bioinformatics and software design knowledge, background comprehension for all sections of this Technical Documentation, and acting to facilitate the use of a deployed TersectBrowser+ by the plant breeder main user of the software. When a new dataset is requested for deployment on TersectBrowser+, the Admin will need to follow the below steps:
\begin{enumerate}
    \item Collect background context and required data from the user.
\begin{itemize}
    \item Reference genome in fasta format (may be compressed). Size and chromosome number will impact the speed of dataset upload and deployment - TersectBrowser+ has been tested with a genome size of ~1GB, and 20 chromosome pairs.
    \item Resequenced genome dataset in VCF format (may be compressed, as a multi-sample VCF or as a directory of individual files). TersectBrowser+ has been tested with a dataset size of ~500 VCF files representing individual accessions. 
    \item Any metadata associated with different accessions in the dataset, such as wild variety vs domesticated variety. \hl{This should be provided in a text file.}
    \item A GFF file of gene model information for the reference genome, produced using SNPeff software. If not using SNPeff, ensure that variant impact levels are categorised into 'High', 'Medium', and 'Low' impact. If not provided, the gene model track in the Variant Browser will be unavailable. 
    \item Whether there are prior identified introgressions known to be functionally relevant for the species. These can be used to 'check' the ability of TersectBrowser+ to identify introgressed regions for the specified dataset.
\end{itemize}
    \item Clone the GitHub repository of TersectBrowser+ to the local machine. Follow the README from \hyperlink {https://github.com/Tersect-Browser/Tersect-browser.git}{TersectBrowser+ GitHub} for installation instructions.
    \item Download relevant files from section (1) to a new folder within the root directory of the local Tersect-browser repository.
    \item Set environment variables \verb+fasta+, \verb+gff+, and \verb+vcfs+ according to their new location relative to the root of Tersect-browser.
    \item Run the script \verb+setup_new_tbrowser_dataset.py+ using the environment variables of files as input:

@@

The method \verb +JbrowserWrapper+ takes as arguments the props from \verb +JbrowseWrapperProps+ and defines the \verb +viewState+. If an accession has been selected and the accession name is stored under \verb +props.location.accession.name+, the \verb +viewState+ defined in the \verb +JbrowseWithAccessionName+ is returned (See below parameters). If not, a conditional check verifies whether the props \verb +defaultInterval+ and \verb +offsetCanvas+ have been populated. If either of them are empty, a container with the phrase "Loading…" is returned. If all data is present, the \verb +viewState+ is defined using the function \verb +createViewState+. This takes as arguments the following variables: 
\begin{itemize}
\item \verb +assembly+: the assembly track 
\item \verb +tracks+: the data tracks to display 
\item \verb +configuration+: the config variable where the theme is defined 
\item \verb +defaultSession+: an object of type LinearGenomeView, which defines the initial state. This object includes the following configurations: 
\begin{itemize}
\item \verb +bpPerPx+: the base pairs per pixel, which defines the bin size 
\item \verb +assemblyName+: the corresponding assembly for the track 
\item \verb +start+: the track start position 
\item \verb +end+: the track end position 
\item \verb +refName+: the genomic coordinates of the viewing window 
\end{itemize}
\end{itemize}

@@

The configurations for the assembly track, which is built upon the reference accession SL2.50, is defined in \verb +assembly.ts+ in json format. The name is set to the reference accession name and \verb +trackId+ is set to "SL2.50-ReferenceSequenceTrack". The paths to the fasta file, along with its corresponding fasta index, are specified as local server URLs provided by the backend during the Tersect Browser setup.

\paragraph{Tracks Config} 
The configurations for the \verb +VariantTrack+ and Gene Models \verb +FeatureTrack+ are defined in \verb +tracks.ts+ in json format. For the variant tracks, the \verb +name+ and \verb +trackId+ are set as the accession name, and for the Gene Models track these are set to "ITAG2.4 Gene Models".  The paths to the zipped files, along with its corresponding Tabix index files, are specified as local server URLs provided by the backend during the Tersect Browser setup. Each track is separated by a comma.

\paragraph{Styling}
The container styling is defined in \verb +jbrowseConfig.ts+. The palette theme is set to colour '\#459e00' and the \verb +boxShadow+ is set to 'none'.

@@

\paragraph{Frontend Display Feature Description}
When the user clicks on a bin, the popup menu appears and the button "Create barcode" is visible, styled with a barcode icon. When the user clicks on this button, a popup window appears in the centre of the screen. The window title is the corresponding accession name. The window contains the header "Barcode Generator" and contains a description of what is being generated. There is a box for the user to input barcode size, with default set to 150, and the maximum number of variants allowed per barcode (this is an optional parameter. Leaving this blank means all possible barcodes will be returned).

At the bottom is a green "Generate" button. Hovering over the button turns it dark green, and clicking on this button will call the backend scripts to generate the barcodes output. Whilst scripts are running and barcodes are being generated, a spinner is added to the button to indicate loading. After the script has been generated, the spinner disappears and another green button "Download" appears at the bottom left of the window. Clicking on the button downloads the tsv file containing the barcodes.

@@

The method \verb+generateBarcode()+ calls and subscribes to the \verb+tersectBackendService generateBarcodes()+ method, setting the \verb+isLoading+ constant to "True" and passing as arguments \verb+modalTitle+, \verb+chromosome+, \verb+startPosition+, \verb+endPosition+, \verb+barcodeSize+, and \verb+maxVariants+. The returned URL and filename are saved to the variables \verb+downloadUrl+ and \verb+downloadFileName+. Once returned, \verb+isLoading+ is set to "False" and \verb+barcodeSize+ is set back to "150".

The popup window is reset when closed by \verb+resetModal()+ on closing, which sets the \verb+barcodeSize+ to "150", \verb+maxVariants+ and \verb+downloadUrl+ to null, and \verb+downloadFileName+ to an empty string, removing the "Download" button from the popup.

\paragraph{Modal Service Styling}:\\
The modal template is defined in \verb+global-barcode.component.html+. The popup window is visible depending on the modal state isVisible, and is centrally position with \verb+modalTitle+ set as the header. Explanation about the barcode service is described in the \verb+modal-title+ and \verb+modal-subtitle+ divs, and the \verb+input-container+ div contains input boxes for barcode size and maximum number of variants, the button "Generate" which calls \verb+generateBarcode()+ and simultaneously shows a \verb+progressSpinner+ whilst the barcodes are being generated. An anchor displays a clickable "Download" p-button that points towards the generated barcode file. The filename is set to \verb+downloadFileName+, and "Download" is only visible once a \verb+downloadUrl+ has been generated for the barcodes file.

@@

The barcodes and respective metrics are written to a tsv file. The file name is formatted to include system date and time, \verb+'TB_Barcode_Gen'+, and the specified accession.

>>>>

\section{Architecture}
The TersectBrowser+ software is an Angular project, with an original frontend and backend written in Angular 8, and a separate extension space written in Angular 19 using Node 22. This `bolt-on' approach was chosen for the extensions in order to seamlessly integrate the new extensions with the original workspace within the project time constrainsts. Additionally, this approach allows the new extnesions to use up-to-date packages, resulting in enhanced functionality and a modern design. To run TersectBrowser+ on different platforms, the new extensions are hosted on NPM and can be linked as dependencies within the project.

@@

The Admin is defined as an individual with bioinformatics and software design knowledge, background comprehension for all sections of this Technical Documentation, and acting to facilitate the use of a deployed TersectBrowser+ by the plant breeder main user of the software. When a new dataset is requested for deployment on TersectBrowser+, the Admin will need to follow the below steps:
\begin{enumerate}
    \item Collect background context and required data from the user.
\begin{itemize}
    \item Reference genome in fasta format (may be compressed). Size and chromosome number will impact the speed of dataset upload and deployment - TersectBrowser+ has been tested with a genome size of ~1GB, and 20 chromosome pairs.
    \item Resequenced genome dataset in VCF format (may be compressed, as a multi-sample VCF or as a directory of individual files). TersectBrowser+ has been tested with a dataset size of ~500 VCF files representing individual accessions. 
    \item Any metadata associated with different accessions in the dataset, such as wild variety vs domesticated variety. \hl{This should be provided in a text file.}
    \item A GFF file of gene model information for the reference genome, produced using SNPeff software. If not using SNPeff, ensure that variant impact levels are categorised into `High', `Medium', and `Low' impact. If not provided, the gene model track in the Variant Browser will be unavailable. 
    \item Whether there are prior identified introgressions known to be functionally relevant for the species. These can be used to `check' the ability of TersectBrowser+ to identify introgressed regions for the specified dataset.
\end{itemize}
    \item Clone the GitHub repository of TersectBrowser+ to the local machine. Follow the README from \hyperlink {https://github.com/Tersect-Browser/Tersect-browser.git}{TersectBrowser+ GitHub} for installation instructions.
    \item Download relevant files from section (1) to a new folder within the root directory of the local Tersect-browser repository.
    \item Set environment variables \verb+fasta+, \verb+gff+, and \verb+vcfs+ according to their new location relative to the root of Tersect-browser.
    \item Run the script \verb+setup_new_tbrowser_dataset.py+ using the environment variables of files as input:

@@

The method \verb +JbrowserWrapper+ takes as arguments the props from \verb +JbrowseWrapperProps+ and defines the \verb +viewState+. If an accession has been selected and the accession name is stored under \verb +props.location.accession.name+, the \verb +viewState+ defined in the \verb +JbrowseWithAccessionName+ is returned (See below parameters). If not, a conditional check verifies whether the props \verb +defaultInterval+ and \verb +offsetCanvas+ have been populated. If either of them are empty, a container with the phrase ``Loading…" is returned. If all data is present, the \verb +viewState+ is defined using the function \verb +createViewState+. This takes as arguments the following variables: 
\begin{itemize}
\item \verb +assembly+: the assembly track 
\item \verb +tracks+: the data tracks to display 
\item \verb +configuration+: the config variable where the theme is defined 
\item \verb +defaultSession+: an object of type LinearGenomeView, which defines the initial state. This object includes the following configurations: 
\begin{itemize}
\item \verb +bpPerPx+: the base pairs per pixel, which defines the bin size 
\item \verb +assemblyName+: the corresponding assembly for the track 
\item \verb +start+: the track start position 
\item \verb +end+: the track end position 
\item \verb +refName+: the genomic coordinates of the viewing window 
\end{itemize}
\end{itemize}

@@

The configurations for the assembly track, which is built upon the reference accession SL2.50, is defined in \verb +assembly.ts+ in json format. The name is set to the reference accession name and \verb +trackId+ is set to ``SL2.50-ReferenceSequenceTrack". The paths to the fasta file, along with its corresponding fasta index, are specified as local server URLs provided by the backend during the Tersect Browser setup.

\paragraph{Tracks Config} 
The configurations for the \verb +VariantTrack+ and Gene Models \verb +FeatureTrack+ are defined in \verb +tracks.ts+ in json format. For the variant tracks, the \verb +name+ and \verb +trackId+ are set as the accession name, and for the Gene Models track these are set to ``ITAG2.4 Gene Models".  The paths to the zipped files, along with its corresponding Tabix index files, are specified as local server URLs provided by the backend during the Tersect Browser setup. Each track is separated by a comma.

\paragraph{Styling}
The container styling is defined in \verb +jbrowseConfig.ts+. The palette theme is set to colour `\#459e00' and the \verb +boxShadow+ is set to `none'.

@@

\paragraph{Frontend Display Feature Description}
When the user clicks on a bin, the popup menu appears and the button ``Create barcode" is visible, styled with a barcode icon. When the user clicks on this button, a popup window appears in the centre of the screen. The window title is the corresponding accession name. The window contains the header ``Barcode Generator" and contains a description of what is being generated. There is a box for the user to input barcode size, with default set to 150, and the maximum number of variants allowed per barcode (this is an optional parameter. Leaving this blank means all possible barcodes will be returned).

At the bottom is a green ``Generate" button. Hovering over the button turns it dark green, and clicking on this button will call the backend scripts to generate the barcodes output. Whilst scripts are running and barcodes are being generated, a spinner is added to the button to indicate loading. After the script has been generated, the spinner disappears and another green button ``Download" appears at the bottom left of the window. Clicking on the button downloads the tsv file containing the barcodes.

@@

The method \verb+generateBarcode()+ calls and subscribes to the \verb+tersectBackendService generateBarcodes()+ method, setting the \verb+isLoading+ constant to ``True" and passing as arguments \verb+modalTitle+, \verb+chromosome+, \verb+startPosition+, \verb+endPosition+, \verb+barcodeSize+, and \verb+maxVariants+. The returned URL and filename are saved to the variables \verb+downloadUrl+ and \verb+downloadFileName+. Once returned, \verb+isLoading+ is set to ``False" and \verb+barcodeSize+ is set back to ``150".

The popup window is reset when closed by \verb+resetModal()+ on closing, which sets the \verb+barcodeSize+ to ``150``, \verb+maxVariants+ and \verb+downloadUrl+ to null, and \verb+downloadFileName+ to an empty string, removing the ``Download" button from the popup.

\paragraph{Modal Service Styling}:\\
The modal template is defined in \verb+global-barcode.component.html+. The popup window is visible depending on the modal state isVisible, and is centrally position with \verb+modalTitle+ set as the header. Explanation about the barcode service is described in the \verb+modal-title+ and \verb+modal-subtitle+ divs, and the \verb+input-container+ div contains input boxes for barcode size and maximum number of variants, the button ``Generate" which calls \verb+generateBarcode()+ and simultaneously shows a \verb+progressSpinner+ whilst the barcodes are being generated. An anchor displays a clickable ``Download" p-button that points towards the generated barcode file. The filename is set to \verb+downloadFileName+, and ``Download" is only visible once a \verb+downloadUrl+ has been generated for the barcodes file.

@@

The barcodes and respective metrics are written to a tsv file. The file name is formatted to include system date and time, \verb+`TB_Barcode_Gen'+, and the specified accession.
